The Etiology of Cybersecurity

### 2020-08-03 reports/paper_0/main.tex
intro review

\section{Introduction}\label{sec:intro}
\paragraph{Context and motivation} 
A \emph{scientific theory} is an explanation of a phenomenon such that the
explanation follows the scientific method. The \emph{scientific method} is an
\emph{empirical} method that aims at mitigating potential fallacies in
theories.  Karl Popper famously argued (e.g. in \autocite{Popper1959logic})
that a scientific theory can never be verified but only falsified, that a
theory should not be conceived by using the principle of
induction\footnote{Albert  Einstein wrote to Popper: ``[...] and I think (like
you, by the way) that theory cannot be fabricated out of the results of
observation, but that it can only be invented.'' \autocite{Popper1959logic}},
and that empirical experiments should be considered as the only evidence to
support the non-falseness of a scientific theory. 
If we consider a natural (real-world) phenomenon, such as gravity, we can think
of a scientific theory as an \emph{abstract} mathematical formula (or a set of
formulas) that defines axiomatically a certain (natural) phenomenon in a
specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof can be carried out to verify
that the theory properly abstracts the phenomenon it describes (i.e. that the
theory is a sound or complete approximation of the real phenomenon).
Similarly, the results of the application of a scientific theory (i.e. its
predictions) can be verified in the abstract mathematical structure but how can
we be sure that the theory correctly characterizes the phenomenon it describes? 
Popper's famous ``demarcation principle'' draws a line between scientific and
pseudo-scientific theories by proposing that a scientific theory can
be considered acceptable as long as no empirical experiment falsify it. For
example, the theory of gravity, in its mathematical setting, can predict how
objects attract each other, and those predictions can be tested by empirical
experiment. If the theory correctly abstracts the phenomenon, no empirical
experiment should falsify the predictions of that theory.

@@

As an illustrative example, in Figure~\ref{fig:protocol-example}, under the perfect
cryptography assumption (i.e. assuming that the cryptographic primitives have
no cybersecurity flaw), and assuming that no violation to any cybersecurity property
occurs after message I), the freedom of choosing the scope determines that the
flaws related to the dishonest impersonification of the Server may or may not
be considered in the verification process.  This choice has a tremendous impact
on the focus and findings of the verification of the cybersecurity of the protocol.
While this may seem to turn upon minutiae and foreseeable, this highlights the
false sense of security that may derive from a non-falsifiable theory of system
cybersecurity. 
In other words, if (as Herley mentions in \autocite{Herley2016usenixvideo}) 
the problem is in the way we reason on cybersecurity, we should
see it in the way we reply to simple cybersecurity (foundamental) questions.
Let us consider the foundamental question ``what is cybersecurity?'',
as in \autocite{hintikka2013inquiry}, we can translate the wh-question into an equivalent
existentially quantified formula. 
But if the formula is $\exists P. P(s)\rightarrow Secure(s)$,
(where we ask if there exists a predicate $P$ such that, given any arbitrary 
system $s$, when $P$ holds for $s$ then $s$ can be considered secure),
when the answer is (for example) ``if the system is confidential, the
system is secure'' the formula becomes tautological. In fact, $P$ and $Secure$ can 
be considered as semantically equivalent (cybersecurity is confidentiality and vice-versa).

@@

we reason on what causes insecurities.
We focus on insecurities only caused by the exploitation of
cybersecurity attacks, and we assume that achieving cybersecurity means preventing all those attacks 
from being exploitable or exploited.
Our hypothesis is that cybersecurity attacks are only caused by 
the presence of errors in the design or implementation of a system (i.e. cybersecurity weaknesses).
For example, a design or implementation weakness in a sanitization function
may allow an attacker to perform SQL-injections. If weaknesses are necessary
to enable cybersecurity attacks, a theory that predicts all potential 
weaknesses can be effectively used to predict the
(in)security of a system.
Therefore, 
a possible answer to the initial question can be that (definition) cybersecurity is the 
absence of attacks and (claim) the absence of weaknesses implies cybersecurity.
This, however, raises another question: ``which are all the errors/weaknesses
in (the engineering of) a system that make it possible for a cybersecurity attack
to happen?''. More rigorously, which scientific theory predicts all
possible weaknesses that causes cybersecurity attacks?

>>>>

\section{Introduction}\label{sec:intro}
\paragraph{Context and motivation} 
A \emph{scientific theory} is an explanation of a phenomenon such that the
explanation follows the scientific method. The \emph{scientific method} is an
\emph{empirical} method that aims at mitigating potential fallacies in
theories.  Karl Popper famously argued (e.g. in \autocite{Popper1959logic})
that a scientific theory can never be verified but only falsified, that a
theory should not be conceived by using the principle of
induction\footnote{Albert  Einstein wrote to Popper: ``[...] and I think (like
you, by the way) that theory cannot be fabricated out of the results of
observation, but that it can only be invented.'' \autocite{Popper1959logic}},
and that empirical experiments should be considered as the only evidence to
support the non-falseness of a scientific theory. 
If we consider a natural (real-world) phenomenon, such as gravity, we can think
of a scientific theory as an \emph{abstract} mathematical formula (or a set of
formulas) that defines axiomatically a certain (natural) phenomenon in a
specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof can be carried out to verify
that the theory properly abstracts the phenomenon it describes (i.e. that the
theory is a sound or complete approximation of the real phenomenon).
Similarly, the results of the application of a scientific theory (i.e. its
predictions) can be verified in the abstract mathematical structure but how can
we be sure that the theory correctly characterizes the phenomenon it describes? 
Popper's famous ``demarcation principle'' draws a line between scientific and
pseudo-scientific theories by proposing that a scientific theory can
be considered acceptable as long as no empirical experiment falsify it. For
example, a theory of gravity, in its mathematical setting, can predict how
objects attract each other, and those predictions can be tested by empirical
experiment. If the theory correctly represents the phenomenon, no empirical
experiment should falsify the predictions of that theory.

@@

As an illustrative example, in Figure~\ref{fig:protocol-example}, under the
perfect cryptography assumption (i.e. assuming that the cryptographic
primitives have no cybersecurity flaw), and assuming that no violation to any
cybersecurity property occurs after message I), the freedom of choosing the
scope determines that the flaws related to the dishonest impersonification of
the Server may or may not be considered in the verification process.  This
choice has a tremendous impact on the focus and findings of the verification of
the cybersecurity of the protocol.  While this may seem to turn upon minutiae
and foreseeable, this highlights the false sense of security that may derive
from a non-falsifiable theory of system cybersecurity.  In other words, as
Herley mentions in \autocite{Herley2016usenixvideo}, the problem is in the way
we reason on cybersecurity, and we can see it in the way we reply to simple
cybersecurity (foundamental) questions.  Let us consider the general question
``what is cybersecurity?''; as in \autocite{hintikka2013inquiry}, we can
translate the wh-question into an equivalent existentially quantified formula.
But if the formula is $\exists P. P(s)\rightarrow Secure(s)$, (where we ask if
there exists a predicate $P$ such that, given any arbitrary system $s$, when
$P$ holds for $s$ then $s$ can be considered secure), when the answer is (for
example) ``if the system is confidential, the system is secure'' the formula
becomes tautological. In fact, $P$ and $Secure$ can be considered as
semantically equivalent (cybersecurity is confidentiality and vice-versa).

@@

we reason on what causes insecurities.  We focus on insecurities only caused by
the exploitation of cybersecurity attacks, and we assume that achieving
cybersecurity means preventing all those attacks from being exploitable or
exploited.  Our hypothesis is that \emph{cybersecurity attacks are only caused by the
presence of errors in the design or implementation of a system} (i.e.
cybersecurity weaknesses).  For example, a design or implementation weakness in
a sanitization function may allow an attacker to perform SQL-injections. If
weaknesses are necessary to enable cybersecurity attacks, a theory that
predicts all potential weaknesses can be effectively used to predict the
(in)security of a system.  Therefore, a possible answer to the initial question
can be that (definition) cybersecurity is the absence of attacks and (claim)
the absence of weaknesses implies cybersecurity.  This, however, raises another
question: ``which are all the errors/weaknesses in (the engineering of) a
system that make it possible for a cybersecurity attack to happen?''. More
rigorously, which scientific theory predicts all possible weaknesses that
causes cybersecurity attacks?
